---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
Load data
```{r}
data_df <- read.csv(unz("activity.zip", "activity.csv"), header = TRUE,
                    na.strings = "NA")
```

Convert date values from factor to date object in R
```{r}
data_df <- transform(data_df, date = as.Date(date))
```

## What is mean total number of steps taken per day?
Compute total number of steps taken per day (ignoring NAs)
```{r}
total_steps_per_day = tapply(data_df$steps, data_df$date, sum, na.rm = TRUE)
```

Histogram of total number of steps per day
```{r}
hist(total_steps_per_day, breaks = 10, xlab = "Total Number of Steps Per Day",
     ylab = "Number of Days",
     main = "Histogram of Total Number of Steps Per Day")
```

Mean of total number of steps taken per day
```{r}
mean(total_steps_per_day, na.rm = TRUE)
```

Median of total number of steps taken per day
```{r}
median(total_steps_per_day, na.rm = TRUE)
```

## What is the average daily activity pattern?
Compute average number of steps for every time interval
```{r}
ave_steps_per_interval = tapply(data_df$steps, data_df$interval,
                                mean, na.rm = TRUE)
```

Time series plot of 5-minute interval and average number of steps taken across all day
```{r}
plot(names(ave_steps_per_interval), ave_steps_per_interval, type = "l",
     xlab = "Interval (min)", ylab = "Average Number of Steps",
     main = "Plot of Average Number of Steps across all Days per Time Interval")
```

Get interval which contains the highest number of steps on average across all days
```{r}
names(which.max(ave_steps_per_interval))
```



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
